---
title: "3.2: Clustering with method HiClimR"
author: "Silvia Ventoruzzo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import packages
```{r}
cran_packages <- c("tidyselect", "tidyverse", "sf", "lwgeom", "nngeo", "fastDummies",
                   "HiClimR", "cluster", "ggdendro", "lubridate", "parallelDist",
                   "dtw", "factoextra", "ggpubr",  "corrplot", "BBmisc", "tseries",
                   "dtwclust", "clValid", 
                   "gridExtra", "grid", "GGally", "ggforce", "gtools", "viridis")
for (package in cran_packages) {
  if(!require(package, character.only = TRUE)) install.packages(package, character.only = TRUE)
  library(package, character.only = TRUE, quietly = TRUE)
}

rm(cran_packages)
rm(package)
```

Load data from previous script
```{r}
load("../RData/3_Data_for_Clustering.RData")
```

Set default ggplot theme
```{r}
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
theme_update(plot.subtitle = element_text(hjust = 0.5))
```


## HiClimR

HiClimR requires a matrix, or a list of matrices, with N rows as spatial elements (here: stations) and M columns as observations (here: days) for each variable.

Check that all matrices have the same size
```{r}
sapply(names(weather_matrix_list),
       function(i) dim(weather_matrix_list[[i]]))
```

### method = "regional"
Since the variable `station_height` is for most stations constant through time, it needs to be removed from the list, because the function `HiClimR` requires multiple values for each row (station).

```{r}
source("../Helpers/hclust_nclust.R")

HiClimR_regional_results <- HiClimR(x           = weather_matrix_list,
                                    method      = "regional",
                                    lon         = station_data$lon,
                                    lat         = station_data$lat,
                                    validClimR  = TRUE,
                                    plot        = FALSE)

hclust_nclust(hclust_resu = HiClimR_regional_results,
               feature_distance_matrix = corr_matrix,
               spatial_distance_matrix = geodist_matrix,
              colors = cividis(n = 2, end = 0.8))
```

The regional linkage method already gives a number of clusters. However, it is also possible to manually cut the tree. Looking at the results from the graphics displayed before, the chosen number of clusters is 6.
```{r}
ggplot(station_polygons) +
  geom_sf(aes(fill=as.factor(HiClimR_regional_results$region)),
          size = 0.3, alpha = 0.6) +
  labs(title="Map of clusters from HiClimR",
       subtitle = "detrended daily data, method = `regional`",
       caption = paste0("Automatic cut at ", 
                        length(HiClimR_regional_results$regionID), " clusters"),
       fill="Region") +
  scale_fill_manual(values = plasma(n = length(HiClimR_regional_results$regionID))) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

k_HclimR_reg <- 6

ggplot(station_polygons) +
  geom_sf(aes(fill=as.factor(cutree(HiClimR_regional_results, k = k_HclimR_reg))),
          size = 0.3, alpha = 0.6) +
  labs(title="Map of clusters from HiClimR",
       subtitle = "detrended daily data, method = `regional`",
       paste0("Manual cut at ", k_HclimR_reg, " clusters"),
       fill="Region") +
  scale_fill_manual(values = plasma(n = k_HclimR_reg)) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

Most of the stations are in Cluster 1, so we check which ones are in Cluster 2.
```{r}
station_data %>%
  mutate(cluster = HiClimR_regional_results$region) %>%
  filter(cluster == 2)
```

### method = "average", contigConst = 1

An alternative to the newly developed regional linkage method is a "standard" linkage method with a contiguity constraint ([0-1]). Since the adjecency of stations is very important, the contiguity constraint will be set to 1.
```{r}
HiClimR_average_results <- HiClimR(x           = weather_matrix_list,
                                method      = "average",
                                contigConst = 1,
                                lon         = station_data$lon,
                                lat         = station_data$lat,
                                validClimR  = TRUE,
                                plot        = FALSE)
hclust_nclust(hclust_result = HiClimR_average_results,
               feature_distance_matrix = corr_matrix,
               spatial_distance_matrix = geodist_matrix,
              colors = cividis(n = 2, end = 0.8))

rm(weather_matrix_list)
```

Map of clusters
```{r}
k_HclimR_avg <- 4
ggplot(station_polygons) +
  geom_sf(aes(fill=as.factor(cutree(HiClimR_average_results, k = k_HclimR_avg))),
          size = 0.3, alpha = 0.6) +
  labs(title="Map of clusters from HiClimR",
       subtitle = "detrended daily data, method = `average` and contigConst = 1",
       fill="Region") +
  scale_fill_manual(values = plasma(n = k_HclimR_avg)) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

## With weekly data

### method = "regional"
Since the variable `station_height` is for most stations constant through time, it needs to be removed from the list, because the function `HiClimR` requires multiple values for each row (station).

```{r}
HiClimR_regional_weekly_results <- HiClimR(x    = weather_weekly_matrix_list,
                                    method      = "regional",
                                    lon         = station_data$lon,
                                    lat         = station_data$lat,
                                    validClimR  = TRUE,
                                    plot        = FALSE)

hclust_nclust(hclust_result = HiClimR_regional_weekly_results,
               feature_distance_matrix = corr_weekly_matrix,
               spatial_distance_matrix = geodist_matrix,
              colors = cividis(n = 2, end = 0.8))
```

As before, the regional linkage method produces its own cut, but looking at the graphics I also chose another value.
```{r}
ggplot(station_polygons) +
  geom_sf(aes(fill=as.factor(HiClimR_regional_weekly_results$region)),
          size = 0.3, alpha = 0.6) +
  labs(title="Map of clusters from HiClimR",
       subtitle = "weekly data, method = `regional`",
       caption = paste0("Automatic cut at ", length(HiClimR_regional_weekly_results$regionID), " clusters"),
       fill="Region") +
    scale_fill_manual(values = plasma(n = length(HiClimR_regional_weekly_results$regionID))) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

k_HclimR_week_reg <- 6
ggplot(station_polygons) +
  geom_sf(aes(fill=as.factor(cutree(HiClimR_regional_weekly_results, k = k_HclimR_week_reg))),
          size = 0.3, alpha = 0.6) +
  labs(title="Map of clusters from HiClimR",
       subtitle = "weekly data, method = `regional`",
       paste0("Manual cut at ", k_HclimR_week_reg, " clusters"),
       fill="Region") +
    scale_fill_manual(values = plasma(n = k_HclimR_week_reg)) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

```{r}
station_data %>%
  mutate(cluster = HiClimR_regional_weekly_results$region) %>%
  filter(cluster == 2)
```

### method = "average", contigConst = 1
```{r}
HiClimR_average_weekly_results <- HiClimR(x           = weather_weekly_matrix_list,
                                method      = "average",
                                contigConst = 1,
                                lon         = station_data$lon,
                                lat         = station_data$lat,
                                validClimR  = TRUE,
                                plot        = FALSE)
hclust_nclust(hclust_result = HiClimR_average_weekly_results,
               feature_distance_matrix = corr_weekly_matrix,
               spatial_distance_matrix = geodist_matrix,
              colors = cividis(n = 2, end = 0.8))

rm(weather_weekly_matrix_list)
```

Map of clusters
```{r}
k_HclimR_week_avg <- 5
ggplot(station_polygons) +
  geom_sf(aes(fill=as.factor(cutree(HiClimR_average_weekly_results, k = k_HclimR_week_avg))),
          size = 0.3, alpha = 0.6) +
  labs(title="Map of clusters from HiClimR",
       subtitle = "weekly data, method = `average` and contigConst = 1",
       fill="Region") +
  theme_void() +
  scale_fill_manual(values = plasma(n = k_HclimR_week_avg)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

```{r}
rm(list = str_subset(string = str_subset(string = ls(), pattern = "results$", negate = TRUE), 
                     pattern = "^k_", negate = TRUE))
save.image("../RData/4_HiClimR_results.RData")
```