---
title: "3.3: Clustering with method ClustGeo"
author: "Silvia Ventoruzzo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import packages
```{r}
# Detach the following packages and ClustGeo
if ("package:raster" %in% search()) detach("package:raster", unload=TRUE)
if ("package:spacetime" %in% search()) detach("package:spacetime", unload=TRUE)
if ("package:sp" %in% search()) detach("package:sp", unload=TRUE)
if ("package:lattice" %in% search()) detach("package:lattice", unload=TRUE)
if ("package:maps" %in% search()) detach("package:maps", unload=TRUE)

cran_packages <- c("tidyselect", "tidyverse", "cluster", "ggdendro", "lubridate", "factoextra", 
                   "ggpubr",  "corrplot", "BBmisc", "clValid", "ClustGeo",
                   "gridExtra", "grid", "GGally", "ggforce", "gtools", "viridis")
for (package in cran_packages) {
  if(!require(package, character.only = TRUE)) install.packages(package, character.only = TRUE)
  library(package, character.only = TRUE, quietly = TRUE)
}

rm(cran_packages)
rm(package)
```

Load data from previous script
```{r}
load("../RData/3_Data_for_Clustering.RData")
```

Set default ggplot theme
```{r}
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
theme_update(plot.subtitle = element_text(hjust = 0.5))
```

# ClustGeo

The clustering function in `ClustGeo` requires two matrices:
- $D_0$: dissimilarities in the "feature space"
- $D_1$: dissimilarities in the "constraint space"

## Correlation distance

#### Choice of number of clusters
To choose the suitable number of clusters, we use hierarchical clustering considering only $D_0$ (feature dissimilarities) and look at the dendogramm.

```{r}
source("../Helpers/hclust_nclust.R")

D0_results <- hclustgeo(D0 = as.dist(corr_matrix), 
                        scale = TRUE, wt = NULL)
hclust_nclust(hclust_result = D0_results,
               feature_distance_matrix = corr_matrix,
               spatial_distance_matrix = geodist_matrix,
              colors = cividis(n = 2, end = 0.8))
```

The chosen number of clusters is 4. Here is the map of the clustering result.
```{r}
k_CG_corr <- 4
ggplot(station_polygons) +
  geom_sf(aes(fill=as.factor(cutree(D0_results, k = k_CG_corr))),
          size = 0.3, alpha = 0.6) +
  labs(title="Map of clusters from GeoClust",
       subtitle = "using only feature dissimilarity",
       fill="Region") +
    scale_fill_manual(values = plasma(n = k_CG_corr)) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

rm(D0_results)
```

## Choice of the mixing parameter
The idea is to perform separate calculations for feature homogeneity and the geographic cohesion of the partitions obtained for a range of different values of $\alpha$ and the number of clusters previously selected.
The practical solution is: conditioning on K and choosing $\alpha$ that best compromises between loss of feature and increase of geographical homogeneity.
```{r}
choice_of_mixing_param <- choicealpha(D0          = as.dist(corr_matrix),
                                      D1          = as.dist(geodist_matrix),
                                      range.alpha = seq(0, 1, 0.1),
                                      K           = k_CG_corr,  # number of clusters
                                      graph       = FALSE)
```

Plot of explained pseudo-inertias for different values of $\alpha$.
```{r}
plot_list <- list()
plot_list[["Q"]] <- choice_of_mixing_param$Q %>%
  as.data.frame() %>%
  mutate(alpha = choice_of_mixing_param$range.alpha) %>%
  pivot_longer(cols = c("Q0", "Q1"), values_to = "Q") %>%
  ggplot(aes(x = alpha, y = Q, color = name)) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("black", "red"),
                     labels = c("based on D0", "based on D1")) +
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(0, 1)) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  theme_bw() +
  labs(title = "Proportion of explained pseudo-inertias",
       subtitle = "") +
  theme(plot.title = element_text(hjust = 0.5, size = 11),
        plot.subtitle = element_text(hjust = 0.5, size = 9),
        axis.title = element_text(size = 9),
        legend.title = element_blank())

plot_list[["Qnorm"]] <- choice_of_mixing_param$Qnorm %>%
  as.data.frame() %>%
  mutate(alpha = choice_of_mixing_param$range.alpha) %>%
  pivot_longer(cols = c("Q0norm", "Q1norm"), values_to = "Qnorm") %>%
  ggplot(aes(x = alpha, y = Qnorm, color = name)) +
  geom_line() +
  geom_point() +
  # annotate(geom = "text", x = 0, y = 1.25, size = 8,
  #          label = paste0("of ", round(choice_of_mixing_param$Q[1,1]*100), "%")) +
  scale_color_manual(values = c("black", "red"),
                     labels = c("based on D0", "based on D1")) +
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  theme_bw() +
  labs(title = "Normal. prop. of explained pseudo-inertias",
       subtitle = paste0("W/ D0 & alpha = 0: ",
                         round(choice_of_mixing_param$Q[1,1]*100), "%; ",
                         "W/ D1 & alpha = 1: ",
                         round(choice_of_mixing_param$Q[length(choice_of_mixing_param$range.alpha),2]*100),
                         "%")) +
  theme(plot.title = element_text(hjust = 0.5, size = 11),
        plot.subtitle = element_text(hjust = 0.5, size = 9),
        axis.title = element_text(size = 9),
        legend.title = element_blank())

arranged_plot <- ggarrange(plotlist = plot_list, ncol = 2, nrow = 1, 
                           common.legend = TRUE, legend = "bottom")

annotate_figure(arranged_plot,
                top = text_grob(
                  paste0("Empirical choice of the mixing parameter for ", 
                         k_CG_corr, " clusters"), size = 12))

rm(plot_list)
rm(arranged_plot)
```

Calculate the decrease/increase of pseudo-inertias with $\alpha$.
```{r}
as.data.frame(choice_of_mixing_param$Q) %>%
  merge(as.data.frame(choice_of_mixing_param$Qnorm), by="row.names") %>%
  rename(alpha = Row.names) %>%
  mutate(Q0_diff = round(Q0norm - first(Q0norm), 4),
         Q1_diff = round((Q1 - first(Q1))/first(Q1), 4))
rm(choice_of_mixing_param)
```


### Clustering with geographical constraint
After choosing the number of clusters k and the mixing parameter $\alpha$, one can finally cluster using also the spatial distance matrix.
```{r}
alpha_corr <- 0
CG_corr_results <- hclustgeo(D0 = as.dist(corr_matrix),
                        D1 = as.dist(geodist_matrix),
                        alpha = alpha_corr,
                        scale = TRUE, wt = NULL)

# Map of clustering result
ggplot(station_polygons) +
  geom_sf(aes(fill=as.factor(cutree(CG_corr_results, k = k_CG_corr))),
          size = 0.3, alpha = 0.6) +
      scale_fill_manual(values = plasma(n = k_CG_corr)) +
  labs(title="Map of clusters from GeoClust with correlation distance",
       subtitle = paste0("alpha = ", alpha_corr, " from choice using k = ", k_CG_corr),
       fill="Region") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

## Euclidean distance

#### Choice of number of clusters
To choose the suitable number of clusters, we use hierarchical clustering considering only $D_0$ (feature dissimilarities) and look at the dendogramm.

```{r}
D0_results <- hclustgeo(D0 = eucl_matrix_dist, 
                        scale = TRUE, wt = NULL)
hclust_nclust(hclust_result = D0_results,
               feature_distance_matrix = as.matrix(eucl_matrix_dist),
               spatial_distance_matrix = geodist_matrix,
              colors = cividis(n = 2, end = 0.8))
```

There is no consensus across all method for the number of clusters. We will use 4 since it is a middle ground.
```{r}
k_CG_eucl <- 4
ggplot(station_polygons) +
  geom_sf(aes(fill=as.factor(cutree(D0_results, k = k_CG_eucl)))) +
  labs(title="Map of clusters from GeoClust",
       subtitle = "using only feature dissimilarity",
       fill="Region") +
  scale_fill_manual(values = plasma(n = k_CG_eucl)) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

rm(D0_results)
```

## Choice of the mixing parameter
The idea is to perform separate calculations for feature homogeneity and the geographic cohesion of the partitions obtained for a range of different values of $\alpha$ and the number of clusters previously selected.
The practical solution is: conditioning on K and choosing $\alpha$ that best compromises between loss of socio-economic and loss of geographical homogeneity.
```{r}
choice_of_mixing_param <- choicealpha(D0          = eucl_matrix_dist,
                                      D1          = as.dist(geodist_matrix),
                                      range.alpha = seq(0, 1, 0.1),
                                      K           = k_CG_eucl,  # number of clusters
                                      graph       = FALSE)
```

Plot of explained pseudo-inertias for different values of $\alpha$.
```{r}
plot_list <- list()
plot_list[["Q"]] <- choice_of_mixing_param$Q %>%
  as.data.frame() %>%
  mutate(alpha = choice_of_mixing_param$range.alpha) %>%
  pivot_longer(cols = c("Q0", "Q1"), values_to = "Q") %>%
  ggplot(aes(x = alpha, y = Q, color = name)) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("black", "red"),
                     labels = c("based on D0", "based on D1")) +
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(0, 1)) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  theme_bw() +
  labs(title = "Proportion of explained pseudo-inertias",
       subtitle = "") +
  theme(plot.title = element_text(hjust = 0.5, size = 11),
        plot.subtitle = element_text(hjust = 0.5, size = 9),
        axis.title = element_text(size = 9),
        legend.title = element_blank())

plot_list[["Qnorm"]] <- choice_of_mixing_param$Qnorm %>%
  as.data.frame() %>%
  mutate(alpha = choice_of_mixing_param$range.alpha) %>%
  pivot_longer(cols = c("Q0norm", "Q1norm"), values_to = "Qnorm") %>%
  ggplot(aes(x = alpha, y = Qnorm, color = name)) +
  geom_line() +
  geom_point() +
  # annotate(geom = "text", x = 0, y = 1.25, size = 8,
  #          label = paste0("of ", round(choice_of_mixing_param$Q[1,1]*100), "%")) +
  scale_color_manual(values = c("black", "red"),
                     labels = c("based on D0", "based on D1")) +
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  theme_bw() +
  labs(title = "Normal. prop. of explained pseudo-inertias",
       subtitle = paste0("W/ D0 & alpha = 0: ",
                         round(choice_of_mixing_param$Q[1,1]*100), "%; ",
                         "W/ D1 & alpha = 1: ",
                         round(choice_of_mixing_param$Q[length(choice_of_mixing_param$range.alpha),2]*100),
                         "%")) +
  theme(plot.title = element_text(hjust = 0.5, size = 11),
        plot.subtitle = element_text(hjust = 0.5, size = 9),
        axis.title = element_text(size = 9),
        legend.title = element_blank())

arranged_plot <- ggarrange(plotlist = plot_list, ncol = 2, nrow = 1, 
                           common.legend = TRUE, legend = "bottom")

annotate_figure(arranged_plot,
                top = text_grob(
                  paste0("Empirical choice of the mixing parameter for ", 
                         k_CG_eucl, " clusters"), size = 12))

rm(plot_list)
rm(arranged_plot)
```

Calculate the decrease/increase of pseudo-inertias with alpha.
```{r}
as.data.frame(choice_of_mixing_param$Q) %>%
  merge(as.data.frame(choice_of_mixing_param$Qnorm), by="row.names") %>%
  rename(alpha = Row.names) %>%
  mutate(Q0_diff = round(Q0norm - first(Q0norm), 4),
         Q1_diff = round((Q1 - first(Q1))/first(Q1), 4))
rm(choice_of_mixing_param)
```


### Clustering with geographical constraint
```{r}
alpha_eucl = 0.4
CG_eucl_results <- hclustgeo(D0 = eucl_matrix_dist,
                        D1 = as.dist(geodist_matrix),
                        alpha = alpha_eucl,
                        scale = TRUE, wt = NULL)
ggdendrogram(CG_eucl_results, rotate = FALSE, size = 2) +
  labs(title = "Dendogramm from GeoClust",
       subtitle = paste0("alpha = ", alpha_eucl, " from choice using k = ", k_CG_eucl)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

ggplot(station_polygons) +
  geom_sf(aes(fill=as.factor(cutree(CG_eucl_results, k = k_CG_eucl))),
          size = 0.3, alpha = 0.6) +
  labs(title="Map of clusters from GeoClust with Euclidean distance",
       subtitle = paste0("alpha = ", alpha_eucl, " from choice using k = ", k_CG_eucl),
       fill="Region") +
    scale_fill_manual(values = plasma(n = k_CG_eucl)) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

## DTW distance

#### Choice of number of clusters
To choose the suitable number of clusters, we use hierarchical clustering considering only $D_0$ (feature dissimilarities) and look at the dendogramm.

```{r}
D0_results <- hclustgeo(D0 = as.dist(dtw_matrix_dist), 
                        scale = TRUE, wt = NULL)
hclust_nclust(hclust_result = D0_results,
               feature_distance_matrix = as.matrix(dtw_matrix_dist),
               spatial_distance_matrix = geodist_matrix,
              colors = cividis(n = 2, end = 0.8))
```

There is no consensus across all method for the number of clusters. We will use 4 since it is a middle ground.
```{r}
k_CG_dtw = 4
ggplot(station_polygons) +
  geom_sf(aes(fill=as.factor(cutree(D0_results, k = k_CG_dtw))), size = 0.5) +
  labs(title="Map of clusters from GeoClust",
       subtitle = "using only feature dissimilarity",
       fill="cluster") +
  scale_fill_manual(values = plasma(n = k_CG_dtw)) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
       
rm(D0_results)
```

## Choice of the mixing parameter
The idea is to perform separate calculations for feature homogeneity and the geographic cohesion of the partitions obtained for a range of different values of $\alpha$ and the number of clusters previously selected.
The practical solution is: conditioning on K and choosing $\alpha$ that best compromises between loss of socio-economic and loss of geographical homogeneity.
```{r}
choice_of_mixing_param <- choicealpha(D0          = as.dist(dtw_matrix_dist),
                                      D1          = as.dist(geodist_matrix),
                                      range.alpha = seq(0, 1, 0.1),
                                      K           = k_CG_dtw,  # number of clusters
                                      graph       = FALSE)
```


```{r}
plot_list <- list()
plot_list[["Q"]] <- choice_of_mixing_param$Q %>%
  as.data.frame() %>%
  mutate(alpha = choice_of_mixing_param$range.alpha) %>%
  pivot_longer(cols = c("Q0", "Q1"), values_to = "Q") %>%
  ggplot(aes(x = alpha, y = Q, color = name)) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("black", "red"),
                     labels = c("based on D0", "based on D1")) +
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(0, 1)) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  theme_bw() +
  labs(title = "Proportion of explained pseudo-inertias",
       subtitle = "") +
  theme(plot.title = element_text(hjust = 0.5, size = 11),
        plot.subtitle = element_text(hjust = 0.5, size = 9),
        axis.title = element_text(size = 9),
        legend.title = element_blank())

plot_list[["Qnorm"]] <- choice_of_mixing_param$Qnorm %>%
  as.data.frame() %>%
  mutate(alpha = choice_of_mixing_param$range.alpha) %>%
  pivot_longer(cols = c("Q0norm", "Q1norm"), values_to = "Qnorm") %>%
  ggplot(aes(x = alpha, y = Qnorm, color = name)) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("black", "red"),
                     labels = c("based on D0", "based on D1")) +
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  theme_bw() +
  labs(title = "Normal. prop. of explained pseudo-inertias",
       subtitle = paste0("W/ D0 & alpha = 0: ",
                         round(choice_of_mixing_param$Q[1,1]*100), "%; ",
                         "W/ D1 & alpha = 1: ",
                         round(choice_of_mixing_param$Q[length(choice_of_mixing_param$range.alpha),2]*100),
                         "%")) +
  theme(plot.title = element_text(hjust = 0.5, size = 11),
        plot.subtitle = element_text(hjust = 0.5, size = 9),
        axis.title = element_text(size = 9),
        legend.title = element_blank())

arranged_plot <- ggarrange(plotlist = plot_list, ncol = 2, nrow = 1, 
                           common.legend = TRUE, legend = "bottom")

annotate_figure(arranged_plot,
                top = text_grob(
                  paste0("Empirical choice of the mixing parameter for ", 
                         k_CG_dtw, " clusters"), size = 12))

rm(plot_list)
rm(arranged_plot)
```

Calculate the decrease/increase of pseudo-inertias with alpha.
```{r}
as.data.frame(choice_of_mixing_param$Q) %>%
  merge(as.data.frame(choice_of_mixing_param$Qnorm), by="row.names") %>%
  rename(alpha = Row.names) %>%
  mutate(Q0_diff = round(Q0norm - first(Q0norm), 4),
         Q1_diff = round((Q1 - first(Q1))/first(Q1), 4))
rm(choice_of_mixing_param)
```


### Clustering with geographical constraint
```{r}
alpha_dtw <- 0.2
CG_dtw_results <- hclustgeo(D0 = as.dist(dtw_matrix_dist),
                        D1 = as.dist(geodist_matrix),
                        alpha = alpha_dtw,
                        scale = TRUE, wt = NULL)
ggdendrogram(CG_dtw_results, rotate = FALSE, size = 2) +
  labs(title = "Dendogramm from GeoClust",
       subtitle = paste0("alpha = ", alpha_dtw, " from choice using k = ", k_CG_dtw)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

ggplot(station_polygons) +
  geom_sf(aes(fill=as.factor(cutree(CG_dtw_results, k = k_CG_dtw))), 
          size = 0.3, alpha = 0.6) +
  labs(title="Map of clusters from GeoClust with DTW distance",
       subtitle = paste0("alpha = ", alpha_dtw, " from choice using k = ", k_CG_dtw),
       fill="Region") +
  scale_fill_manual(values = plasma(n = k_CG_dtw)) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

```{r}
rm(list = str_subset(string = str_subset(string = str_subset(string = ls(), 
                                                             pattern = "^alpha", negate = TRUE), 
                                         pattern = "results$", negate = TRUE), 
                     pattern = "^k_", negate = TRUE))
save.image("../RData/5_ClustGeo_results.RData")
```