---
title: "3.6: Comparison of Clustering Methods"
author: "Silvia Ventoruzzo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import packages
```{r}
# Detach the following packages and ClustGeo
if ("package:raster" %in% search()) detach("package:raster", unload=TRUE)
if ("package:spacetime" %in% search()) detach("package:spacetime", unload=TRUE)
if ("package:sp" %in% search()) detach("package:sp", unload=TRUE)
if ("package:lattice" %in% search()) detach("package:lattice", unload=TRUE)
if ("package:maps" %in% search()) detach("package:maps", unload=TRUE)

cran_packages <- c("tidyselect", "tidyverse", "cluster", "ggdendro", "lubridate", "factoextra", 
                   "ggpubr",  "corrplot", "BBmisc", "clValid", "ClustGeo",
                   "gridExtra", "grid", "GGally", "ggforce", "gtools", "viridis")
for (package in cran_packages) {
  if(!require(package, character.only = TRUE)) install.packages(package, character.only = TRUE)
  library(package, character.only = TRUE, quietly = TRUE)
}

rm(cran_packages)
rm(package)
```

Load data from previous script
```{r}
load("../RData/3_Data_for_Clustering.RData")
load("../RData/4_HiClimR_results.RData")
load("../RData/5_ClustGeo_results.RData")
load("../RData/6_Bootstrap_ClustGeo_results.RData")
load("../RData/7_CorClustST_results.RData")
```

Set default ggplot theme
```{r}
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
theme_update(plot.subtitle = element_text(hjust = 0.5))
```

## Comparison of clustering results

### Internal validation statistics
Dunn index: should be maximized
Average distance within clusters: should be minimized
Average silhouette width: should be maximized
Connectivity: should be minimized
```{r}
source("../Helpers/spatialcluster_comparison.R")

corr_comparisons <- spatialcluster_comparison(
  labels_list = list(HiClimR_regional_daily = cutree(HiClimR_regional_results, k = k_HclimR_reg),
                     HiClimR_average_daily  = cutree(HiClimR_average_results, k = k_HclimR_avg),
                     CorClustST = CorClustST_results$clusters$label),
  feature_distance_matrix = corr_matrix, 
  spatial_distance_matrix = geodist_matrix,
  stats = c("dunn", "average.within", 
            "avg.silwidth",
            "connectivity"))

corr_weekly_comparisons <- spatialcluster_comparison(
  labels_list = list(HiClimR_regional_weekly = cutree(HiClimR_regional_weekly_results, k = k_HclimR_week_reg),
                     HiClimR_average_weekly = cutree(HiClimR_average_weekly_results, k = k_HclimR_week_avg)),
  feature_distance_matrix = corr_weekly_matrix, 
  spatial_distance_matrix = geodist_matrix,
  stats = c("dunn", "average.within", 
            "avg.silwidth",
            "connectivity"))

eucl_comparisons <- spatialcluster_comparison(
  labels_list = list(CG_euclidean = cutree(CG_eucl_results, k = k_CG_eucl),
                     BCG  = cutree(BCG_results, k = k_CG_eucl)),
  feature_distance_matrix = as.matrix(eucl_matrix_dist), 
  spatial_distance_matrix = geodist_matrix,
  stats = c("dunn", "average.within", 
            "avg.silwidth",
            "connectivity"))

bind_rows(list(corr_comparisons$validation_statistics,
               eucl_comparisons$validation_statistics,
               corr_weekly_comparisons$validation_statistics)) %>%
  arrange(distance, cluster_method)

rm(list=lsf.str())
```

According to the internal validation techniques above, the best performing methods is ClustGeo (with Euclidean distance).

```{r}
# Assignment of cluster info
station_data <- station_data %>%
  mutate(cluster = cutree(CG_eucl_results, k = k_CG_eucl))
station_points <- station_points %>%
  mutate(cluster = cutree(CG_eucl_results, k = k_CG_eucl))
station_polygons <- station_polygons %>%
  mutate(cluster = cutree(CG_eucl_results, k = k_CG_eucl))
weather_data <- weather_data %>%
  left_join(station_data[c("station_id", "cluster")], by = "station_id")
weather_data_num <- weather_data_num %>%
  left_join(station_data[c("station_id", "cluster")], by = "station_id")
weather_data_norm <- weather_data_norm %>%
  left_join(station_data[c("station_id", "cluster")], by = "station_id")
```


```{r}
rm(list = str_subset(string = str_subset(string = str_subset(string = setdiff(ls(), 
                                                                              c("postal_codes_nn",
                                                                              "federal_state_polygons",
                                                                              "postal_code_polygons")), 
                                                             pattern = "^.*variables.*$", negate = TRUE), 
                                         pattern = "^weather", negate = TRUE), 
                     pattern = "^station", negate = TRUE))
save.image("../RData/8_Cluster_Methods_comparison.RData")
```