---
title: "3.7: Analysis of Weather Regions"
author: "Silvia Ventoruzzo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import packages
```{r}
# Detach the following packages and ClustGeo
if ("package:raster" %in% search()) detach("package:raster", unload=TRUE)
if ("package:spacetime" %in% search()) detach("package:spacetime", unload=TRUE)
if ("package:sp" %in% search()) detach("package:sp", unload=TRUE)
if ("package:lattice" %in% search()) detach("package:lattice", unload=TRUE)
if ("package:maps" %in% search()) detach("package:maps", unload=TRUE)

cran_packages <- c("tidyselect", "tidyverse", "cluster", "ggdendro", "lubridate", "factoextra", 
                   "ggpubr",  "corrplot", "BBmisc", "clValid", "ClustGeo",
                   "gridExtra", "grid", "GGally", "ggforce", "gtools", "viridis")
for (package in cran_packages) {
  if(!require(package, character.only = TRUE)) install.packages(package, character.only = TRUE)
  library(package, character.only = TRUE, quietly = TRUE)
}

rm(cran_packages)
rm(package)
```

Load data from previous script
```{r}
load("../RData/8_Cluster_Methods_comparison.RData")
```

Set default ggplot theme
```{r}
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
theme_update(plot.subtitle = element_text(hjust = 0.5))
```


## Analysis of clusters

### Definition of cluster centroids
To improve speed in a future automatization, we will choose as centroid the weather station which has the lowest distance to the cluster center, i.e. the time series of daily means.

```{r}
apply(weather_data_norm, 2, function(x) sum(is.na(x))/length(x))
```


```{r}
# Cluster centers
cluster_centers <- weather_data_norm %>%
  group_by(cluster, measurement_date) %>%
  summarize_if(is.numeric, mean) %>%
  ungroup() %>%
  mutate(station_id = "center")
# Station closest to respective cluster center
# Out of those which did not have missing values in the numeric variables of interest
nearest_to_center <- 
  sapply(unique(station_data$cluster),
         function(j) {
           center <- cluster_centers %>% 
             filter(cluster == j) %>% 
             dplyr::select(-cluster)
           # Stations which did not have missing values in the numeric variables of interest
           points <- weather_data %>%
             filter(cluster == j) %>%
             group_by(station_id) %>%
             summarize_at(vars(paste0(weather_variables, "_missing")), sum) %>%
             filter_at(vars(paste0(weather_variables, "_missing")), function(x) x == 0) %>%
             pull(station_id)
           points <- c(points, "center")
           # Matrices of station data and centers
           data <- 
              sapply(points,
                     function(i) {
                       weather_data_norm %>%
                         bind_rows(center) %>%
                         dplyr::filter(station_id == i) %>%
                         pivot_longer(cols = all_of(numeric_variables), names_to = "variable") %>%
                         pivot_wider(names_from = "measurement_date", 
                                   values_from = "value") %>%
                         dplyr::select(-station_id, -variable, -cluster) %>%
                         as.matrix()
                     },
                   simplify = FALSE, USE.NAMES = TRUE)
           # Calculate distance
           distance <- parDist(x = data, method = "euclidean")
           distance <- as.matrix(distance)
           rownames(distance) <- points
           colnames(distance) <- points
           # Extract point with lowest distance
           nearest_point <- apply(distance, 2, 
                                  function(x) {
                                    x <- x[x > 0]
                                    return(data.frame(nearest = which.min(x),
                                                      distance = min(x),
                                                      stringsAsFactors = FALSE))
                                  })
           return(data.frame(nearest_point = points[nearest_point[["center"]]$nearest],
                             distance = nearest_point[["center"]]$distance,
                             stringsAsFactors = FALSE))
         },
         simplify = FALSE, USE.NAMES = TRUE) %>%
  bind_rows() %>%
  rownames_to_column(var = "cluster") %>%
  mutate(cluster_center = as.integer(cluster))

# Assignment of clusters and centroid info
station_data <- station_data %>%
  left_join(nearest_to_center[c("nearest_point", "cluster_center")], by = c("station_id" = "nearest_point"))
station_points <- station_points %>%
  left_join(nearest_to_center[c("nearest_point", "cluster_center")], by = c("station_id" = "nearest_point"))
station_polygons <- station_polygons %>%
  left_join(nearest_to_center[c("nearest_point", "cluster_center")], by = c("station_id" = "nearest_point"))
weather_data <- weather_data %>%
  left_join(station_data[c("station_id", "cluster_center")], by = "station_id")
weather_data_num <- weather_data_num %>%
  left_join(station_data[c("station_id", "cluster_center")], by = "station_id")
weather_data_norm <- weather_data_norm %>%
  left_join(station_data[c("station_id", "cluster_center")], by = "station_id")
         
# Map of clusters with centroids
station_data %>%
  drop_na(cluster_center) %>%
  ggplot(aes(fill = as.character(cluster))) +
  geom_sf(data = station_polygons, mapping = aes(geometry = polygon), 
          alpha = 0.6, size = 0.3) +
  geom_point(aes(x = lon, y = lat), size = 2) +
  geom_label(aes(x = lon, y = lat+0.3, label = station_name), size = 3, color = "white") +
  scale_fill_manual(values = plasma(n = length(unique(station_data$cluster_center)))) +
  theme_void() +
  labs(title = "Map of regions with respective centroid",
       fill = "Region") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```


Centroid names
```{r}
station_data %>%
  drop_na()
```

### Cluster plausibility
\textbf{Cluster cardinality}: Cluster cardinality is the number of examples per cluster. Plot the cluster cardinality for all clusters and investigate clusters that are major outliers.
\textbf{Cluster magnitude}: Cluster magnitude is the sum of distances from all examples to the centroid of the cluster. Similar to cardinality, check how the magnitude varies across the clusters, and investigate anomalies.
\textbf{Magnitude vs. Cardinality}: Notice that a higher cluster cardinality tends to result in a higher cluster magnitude, which intuitively makes sense. Clusters are anomalous when cardinality doesn't correlate with magnitude relative to the other clusters. Find anomalous clusters by plotting magnitude against cardinality.
```{r}
# Cluster cardinality
ggplot(data = station_data, aes(x = cluster)) +
  geom_bar(fill = "steelblue") +
  theme_bw() +
  labs(title = "Cluster cardinality",
       x = "cluster",
       y = "Points in cluster") +
    theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

# Cluster magnitude
distance_to_centroids <- lapply(
  unique(station_data$cluster),
  function(i) {
    centroid <- station_data %>%
      filter(cluster_center == i) %>%
      pull(station_id)
    points <- station_data %>%
      filter(cluster == i) %>%
      pull(station_id)
    data <- sapply(points,
                   function(k) {
                     weather_data_norm %>%
                       filter(station_id == k) %>%
                       pivot_longer(cols = all_of(numeric_variables), names_to = "variable") %>%
                       pivot_wider(id_cols = c("station_id", "variable"),
                                   names_from = "measurement_date", values_from = "value") %>%
                       dplyr::select(-station_id, -variable) %>%
                       as.matrix()
                     },
                   simplify = FALSE, USE.NAMES = TRUE)
    # Calculate distance
    distance <- parDist(x = data, method = "euclidean")
    distance <- as.matrix(distance)
    rownames(distance) <- points
    colnames(distance) <- points
    return(sum(distance[centroid,]))
  }) %>%
  set_names(nm = unique(station_data$cluster)) %>%
  bind_rows() %>%
  pivot_longer(cols = everything(), names_to = "cluster", values_to = "distance") %>%
  mutate(cluster = as.integer(cluster))

ggplot(data = distance_to_centroids, aes(x = cluster)) +
  geom_bar(aes(y = distance), stat = "identity", fill = "steelblue") +
  labs(title = "Cluster Magnitude",
       x = "cluster",
       y = "Total Point-to-Centroid Distance")

# Magnitude vs. Cardinality
station_data %>%
  group_by(cluster) %>%
  dplyr::count() %>%
  full_join(distance_to_centroids, by = "cluster") %>%
  ggplot(aes(x = n, y = distance)) +
  geom_point(color = "steelblue") +
  geom_smooth(method = lm) +
  geom_text(aes(label = cluster), hjust = 0, vjust = -1) +
  labs(title = "Cardinality vs Magnitude",
       x = "cardinality",
       y = "magnitude")
```

Calculate area size to check that it's not that different despite lower amount of points in cluster 4.
```{r}
cluster_polygons <- station_polygons %>%
  st_set_precision(1e4) %>%
  st_make_valid() %>%
  st_buffer(0.0) %>%
  group_by(cluster) %>%
  summarize(do_union = TRUE) %>%
  ungroup()

cluster_polygons %>%
  mutate(area = st_area(.) %>% units::set_units(value = "km^2")) 
```

Plot cardinality vs area
```{r}
cardinality <- station_data %>%
  group_by(cluster) %>%
  dplyr::count()

cluster_polygons %>%
  mutate(area = st_area(.) %>% units::set_units(value = "km^2")) %>%
  left_join(cardinality, by = "cluster") %>%
  ggplot(aes(x = n, y = area)) +
  geom_point(color = "steelblue") +
  geom_smooth(method = lm) +
  geom_text(aes(label = cluster), hjust = 0, vjust = -1) +
  theme_bw() +
  labs(title = "Cardinality vs Area",
       x = "cardinality",
       y = "area") +
    theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

rm(cardinality)
```

### Plots

Time series of clusters
```{r}
weather_data %>%
  filter(year(measurement_date) == 2015) %>%
  mutate(cluster_center = ifelse(is.na(cluster_center), "point", "center"),
         cluster = paste0("cluster ", cluster)) %>%
  pivot_longer(cols = all_of(weather_variables), names_to = "variable") %>%
  ggplot() +
  geom_line(aes(x = measurement_date, y = value, alpha = cluster_center, color = cluster_center)) +
  facet_grid(variable~cluster, scales = "free_y") +
  theme_bw() +
  labs(color = "", alpha = "") +
  scale_alpha_discrete(range = c(1, 0.5)) +
  scale_color_manual(values = c("red", "grey")) +
  scale_x_date(date_breaks = "3 months",
               date_labels = "%b") +
  theme(legend.position = "bottom") +
  labs(title = "Time series of points and centroids",
       subtitle = "(Year 2015)",
       x = "") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        legend.margin = margin(0,0,0,0),
        legend.box.margin = margin(-10,-10,-10,-10))

# One for each variable
for (v in weather_variables) {
  p <- weather_data %>%
  filter(year(measurement_date) == 2015) %>%
  mutate(cluster_center = ifelse(is.na(cluster_center), "point", "center"),
         cluster = paste0("cluster ", cluster)) %>%
  ggplot() +
  geom_line(aes(x = measurement_date, y = !!sym(v), alpha = cluster_center, color = cluster_center)) +
  facet_wrap(~cluster) +
  theme_bw() +
  labs(color = "", alpha = "") +
  scale_alpha_discrete(range = c(1, 0.5)) +
  scale_color_manual(values = c("red", "grey")) +
  scale_x_date(date_breaks = "3 months",
               date_labels = "%b") +
  theme(legend.position = "bottom") +
  labs(title = paste0("Time series of points and centroids of variable ", v),
       subtitle = "(Year 2015)",
       x = "") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        legend.margin = margin(0,0,0,0),
        legend.box.margin = margin(-10,-10,-10,-10))
  print(p)
}
```

## Keep only centroids

Create table with postal_code + cluster + cluster centroid
```{r}
station_and_centroids <-station_data %>%
  left_join(drop_na(station_data, cluster_center), by = "cluster", suffix = c("", "_centroid")) %>%
  dplyr::select(station_id, cluster, station_id_centroid)

postal_codes_nn <- postal_codes_nn %>%
  dplyr::select(-distance_km) %>%
  left_join(station_and_centroids, by = "station_id")

rm(station_and_centroids)
```

Weather data for centroids
```{r}
weather_data_centroids <- weather_data %>%
  drop_na(cluster_center) %>%
  dplyr::select(cluster, cluster_center, station_id, measurement_date, 
                all_of(variables_of_interest), 
                all_of(paste0(variables_of_interest, "_missing")),
                all_of(paste0(weather_variables, "_score")),
                weather_index)

# Check that there is no date missing for any cluster centroid
weather_data_centroids %>%
  group_by(cluster, station_id) %>%
  dplyr::count()
weather_data_centroids %>%
  group_by(cluster, station_id) %>%
  summarize(n_distinct = n_distinct(measurement_date))
```

### Plots

Boxplots of monthly weather
```{r}
# Boxplot of monthly weather by month
for (m in unique(month(weather_data_centroids$measurement_date, label = TRUE, abbr = FALSE))) {
  p <- weather_data_centroids %>%
    filter(month(measurement_date, label = TRUE, abbr = FALSE) == m) %>%
    pivot_longer(cols = weather_variables, names_to = "variable") %>%
    ggplot(aes(x = as.character(cluster), y = value, fill = as.character(cluster))) +
    geom_boxplot(alpha = 0.5) +
    facet_wrap(~variable, scales = "free_y") +
    scale_fill_manual(values = plasma(n = length(unique(station_data$cluster_center)))) +
    labs(title = "Boxplots of weather variables across region points",
         subtitle = paste0("Month = ", m),
         x = "region") +
      theme(axis.title.y = element_blank(),
            legend.position = "none")
  print(p)
}

# Boxplot of monthly weather by cluster
for (c in unique(station_data$cluster)) {
  p <- weather_data_centroids %>%
    filter(cluster == c) %>%
    mutate(month = month(measurement_date, label = TRUE, abbr = TRUE)) %>%
    pivot_longer(cols = weather_variables, names_to = "variable") %>%
    ggplot(aes(x = month, y = value, fill = month)) +
    geom_boxplot(alpha = 0.5) +
    scale_fill_manual(values = viridis(n = 12)) +
    facet_wrap(~variable, scales = "free_y") +
    labs(title = "Boxplots of weather variables across region points",
         subtitle = paste0("Region = ", c),
         x = "month") +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5),
            axis.title.y = element_blank(),
            legend.position = "none")
  print(p)
}

# Boxplot of monthly weather by variable
for (v in weather_variables) {
  p <- weather_data_centroids %>%
    mutate(month = month(measurement_date, label = TRUE, abbr = FALSE)) %>%
    ggplot(aes(x = as.character(cluster), y = !!sym(v), fill = as.character(cluster))) +
    geom_boxplot(alpha = 0.5) +
    facet_wrap(~month, scales = "free_y") +
    labs(title = paste("Boxplots of monthly", v, "by region"),
         x = "region") +
      theme(axis.title.y = element_blank(),
            legend.position = "none")
  print(p)
}

# Boxplot of yearly weather
weather_data_centroids %>%
  pivot_longer(cols = weather_variables, names_to = "variable") %>%
  ggplot(aes(x = as.character(cluster), y = value, fill = as.character(cluster))) +
  geom_boxplot(alpha = 0.5) +
  facet_wrap(~variable, scales = "free_y") +
  scale_fill_manual(values = plasma(n = length(unique(station_data$cluster_center)))) +
  labs(title = "Distribution of weather variables across region centroids",
       subtitle = "yearly",
       x = "region") +
  theme(axis.title.y = element_blank(),
        legend.position = "none")
```


Line plots
```{r}
# Of specific variable (weekly)
for (v in weather_variables) {
  p <- weather_data_centroids %>%
    mutate(week = week(measurement_date)) %>%
    group_by(cluster, week) %>%
    summarize_at(vars(weather_variables), mean) %>%
    ungroup() %>%
    ggplot(aes(x = week, y = !!sym(v), color = as.character(cluster))) +
    geom_line() +
    scale_x_continuous(breaks = seq(0, 53, 10), minor_breaks = seq(0, 53, 5)) +
    scale_color_manual(values = plasma(n = length(unique(station_data$cluster_center)))) +
    labs(title = "Time series of weekly means across all years",
         subtitle = paste("Variable:", v),
         y = "weekly mean",
         color = "Region") +
    theme(legend.position = "bottom")
  print(p)
}

# Of specific variable (monthly)
for (v in weather_variables) {
  p <- weather_data_centroids %>%
    mutate(month = month(measurement_date, label = FALSE)) %>%
    group_by(cluster, month) %>%
    summarize_at(vars(weather_variables), mean) %>%
    ungroup() %>%
    ggplot(aes(x = month, y = !!sym(v), color = as.character(cluster))) +
    geom_line() +
    scale_x_continuous(breaks = seq(1, 12, 1)) +
    scale_color_manual(values = plasma(n = length(unique(station_data$cluster_center)))) +
    labs(title = "Time series of monthly means across all years",
         subtitle = paste("Variable: ", v),
         y = "monthly mean",
         color = "Region") +
    theme(legend.position = "bottom")
  print(p)
}
```

Comparison of weekly means/medians
```{r}
weather_data_centroids %>%
  group_by(cluster, week = week(measurement_date)) %>%
  summarize_at(vars(weather_variables), mean) %>%
  ungroup() %>%
  pivot_longer(cols = weather_variables, names_to = "variable") %>%
  ggplot(aes(x = as.character(cluster), y = value)) +
  geom_boxplot() +
  facet_wrap(~variable, scales = "free_y") +
  labs(title = "Boxplot of weekly means",
       x = "cluster") +
    theme(axis.title.y = element_blank())

weather_data_centroids %>%
  group_by(cluster, week = week(measurement_date)) %>%
  summarize_at(vars(weather_variables), median) %>%
  ungroup() %>%
  pivot_longer(cols = weather_variables, names_to = "variable") %>%
  ggplot(aes(x = as.character(cluster), y = value)) +
  geom_boxplot() +
  facet_wrap(~variable, scales = "free_y") +
  theme_bw() +
  labs(title = "Boxplot of weekly medians",
       x = "cluster")  +
    theme(axis.title.y = element_blank())
```


Plots of precipitation_type
```{r}
# Bar plot of yearly data
weather_data_centroids %>%
  mutate(cluster = paste0("region ", cluster)) %>%
  ggplot(aes(x = precipitation_type)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..), group = cluster), 
           stat="count", color = "black",
           alpha = 0.8) + 
  facet_wrap(~cluster) +
  scale_y_continuous(labels=scales::percent) +
  labs(title = "Bar plot of precipitation type",
       y = "proportion") +
  theme(axis.title.x = element_blank(),
        legend.position = "none") +
  scale_fill_manual(values = cividis(n = length(unique(weather_data_centroids$precipitation_type))))

# Month vs cluster
weather_data_centroids %>%
  mutate(cluster = paste0("region ", cluster)) %>%
  group_by(cluster, month = month(measurement_date, label = TRUE), precipitation_type) %>%
  dplyr::count() %>%
  ungroup() %>%
  group_by(cluster, month) %>%
  mutate(tot = sum(n)) %>%
  ungroup() %>%
  mutate(percent = n/tot) %>%
  ggplot(aes(x = precipitation_type, y = percent, fill = precipitation_type)) +
  geom_bar(stat = "identity", color = "black", alpha = 0.8) +
  facet_grid(cluster~month) +
  labs(title = "Bar plot of precipitation type by month",
       y = "proportion") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        axis.title.x = element_blank(),
        axis.text = element_text(size = 8),
        legend.position = "none") +
  scale_y_continuous(labels=scales::percent) +
  scale_fill_manual(values = cividis(n = length(unique(weather_data_centroids$precipitation_type))))
```

### Tables

General information about clusters
```{r}
postal_code_inhabitants <- read_csv("../Data/inhabitants_information/plz_einwohner.csv")

postal_code_inhabitants <- postal_code_inhabitants %>%
  group_by(postal_code = str_sub(plz, 1, 3)) %>%
  dplyr::summarize(inhabitants = sum(einwohner)) %>%
  ungroup()

postal_code_inhabitants %>%
  full_join(postal_codes_nn, by = "postal_code") %>%
  group_by(station_id) %>%
  summarize(inhabitants = sum(inhabitants)) %>%
  ungroup() %>%
  full_join(station_polygons, by = "station_id") %>%
  mutate(area = st_area(polygon) %>% units::set_units(value = "km^2")) %>%
  group_by(cluster) %>%
  dplyr::summarize(inhabitants = sum(inhabitants),
                   area        = sum(area),
                   cardinality = n()) %>%
  ungroup()
```


Yearly descriptive statistics
```{r}
source("../Helpers/descriptive_statistics.R")
# Numeric variables
clusters_num <- weather_data_centroids %>%
  group_by(cluster) %>%
  group_map(~descr_numerical(.x))
names(clusters_num)  <- as.character(1:4)
clusters_num <- lapply(names(clusters_num),
               function(i) {
                 clusters_num[[i]] %>%
                   mutate(cluster = i)
               }) %>%
  bind_rows() 

# Categorical variables
clusters_cat <- weather_data_centroids %>%
  dplyr::select(-station_id) %>%
  group_by(cluster) %>%
  group_map(~descr_categorical(.x))
names(clusters_cat)  <- as.character(1:4)
clusters_cat <- lapply(names(clusters_cat),
               function(i) {
                 clusters_cat[[i]] %>%
                   mutate(cluster = i)
               }) %>%
  bind_rows() 

clusters_num %>%
  dplyr::filter(variable %in% variables_of_interest) %>%
  arrange(variable, cluster)
clusters_cat %>%
  dplyr::filter(variable %in% variables_of_interest) %>%
  arrange(factor, cluster) 

rm(list=lsf.str())
```

Monthly descriptive statistics
```{r}
source("../Helpers/descriptive_statistics.R")
# Numeric variables
clusters_monthly_num <- weather_data_centroids %>%
  group_by(cluster, month = lubridate::month(x = measurement_date)) %>%
  group_map(~descr_numerical(.x))
names(clusters_monthly_num) <- expand.grid(cluster = 1:4, month = 1:12) %>%
  mutate(names = paste0(cluster, ";", month)) %>%
  pull(names)
clusters_monthly_num <- lapply(names(clusters_monthly_num),
               function(i) {
                 clusters_monthly_num[[i]] %>%
                   mutate(name = i) %>%
                   separate(col = "name", into = c("cluster", "month"), remove = TRUE)
               }) %>%
  bind_rows() 

# Categorical variables
clusters_monthly_cat <- weather_data_centroids %>%
  dplyr::select(-station_id) %>%
  group_by(cluster, month = lubridate::month(x = measurement_date)) %>%
  group_map(~descr_categorical(.x))
names(clusters_monthly_cat) <- expand.grid(cluster = 1:4, month = 1:12) %>%
  mutate(names = paste0(cluster, ";", month)) %>%
  pull(names)
clusters_monthly_cat <- lapply(names(clusters_monthly_cat),
               function(i) {
                 clusters_monthly_cat[[i]] %>%
                   mutate(name = i) %>%
                   separate(col = "name", into = c("cluster", "month"), remove = TRUE)
               }) %>%
  bind_rows() 

clusters_monthly_num %>%
  filter(variable %in% weather_variables) %>%
  arrange(variable, month, cluster) 
clusters_monthly_cat %>%
  filter(variable %in% variables_of_interest) %>%
  arrange(factor, month, cluster)

rm(list=lsf.str())
```

Remove all except useful objects
```{r}
rm(list = str_subset(string = setdiff(ls(), c("cluster_polygons", "postal_codes_nn", 
                                              "federal_state_polygons", "weather_data", 
                                              "weather_data_centroids", "weather_variables",
                                              "postal_code_polygons", "postal_code_inhabitants")),
                     pattern = "^station",
                     negate = TRUE))
save.image("../RData/9_Weather_Regions.RData")
```
